### 2023-12-13 ITP draft/draft.tex
notes on itp draft

\documentclass[]{article}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{adjustbox}

\newcommand{\tempty}[1]{\llbracket #1 \rrbracket}
\newcommand{\conjM}{\bigwedge\nolimits_M}
\newcommand{\Real}{\mathbb{R}}

%opening
\title{Placeholder title: Formal methods in constraint-based learning}
\author{}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

\begin{itemize}
	\item formal methods in combination with machine learning as approach
	\item normally for verification
	\item use it to deepen understanding/inform ML research itself
\end{itemize}

\section{Background}
\begin{itemize}
	\item A good explanation of DL and similiar methods in machine learning. Including syntax, semantics in tables.
	\item Possibly properties here, not later.
	\item Other work on similiar/related topics (pointing out most formalise finalised trained models, not methods of training itself so can't find a good comparison)
\end{itemize}

\textbf{Some of other work (more or less related depending on the specific case:}

 Formalization of an embedding of linear temporal logic and an associated evaluation function to prove its soundness in Isabelle and producing OCaml versions of the aforementioned functions that can be integrated with Python and PyTorch \cite{chevallier2022constrained}.  Formalisation of other types of algorithms~\cite{daukantas2021trimming} or formalisation of neural networks using Haskell~\cite{xie2023haskell}

 Reviews of works in formal methods in machine learning (not only neural networks) \cite{urban2021review,9842406}. These are however mostly verification work.

 Generally most work is done on trained models - not on the method itself - with very few exceptions. Need to find more of the exceptions, put in related work with caveat that they are not always very directly related to this.

\section{Temporary Name: Formalisation}

Different functions, lemmas, and things that peraphs stand out.

\subsection{Syntax and semantics}

Implementation of the syntax: $expr$ with it's custom dependant type $simple\_type$. Since the type system prevents non-termination of the lambda expressions they, together with let statements, have been omitted in the Coq formalisation (argumentation why it's okay goes here).

Implementation of the semantics: 
\begin{itemize}
	\item $type\_translation$ semantics of types
	\item a word on translation of indices and vectors using tuples
	\item mention Real library from Mathcomp to handle reals
	\item $translation$ - a semantics for all fuzzy DLs. Here there needs to be explanation maybe of how DL2 and STL are separate from the fuzzy logics as they are different enough to warrant a separate definition and lemmas, not enough overlap
	translation itself is dependently typed, handles 4 different fuzzy logics through use of global variable (for now it's global variable, up for discussion)

\end{itemize}

No quantifiers in the semantics. That does not impact the formalisation of most properties aside from soundness as the other properties do not concern quantifiers.

\subsection{Soundness (and associated lemmas)}

\begin{itemize}
	\item $expr\_ind'$ custom induction principle was needed due to use of sequences. Explain sequences in MathComp (hidden tuples)? Explain that this is needed due to possibility of non-associative binary operations in the language.
	\item $[name of DL]\_translate\_Bool\_T\_01$ - lemmas that prove that the fuzzy DL semantics fall within the range [0,1]
\end{itemize}

Inversion lemmas also mentioned here possibly.

\subsection{Logical properties}

On associativity and commutativity for the DLs that have them (and that associativity is much more tricky then commutativity to prove). Not entirely sure if antyhing more interesting is here specifically so likely not its own subsection.

\subsection{Shadow-lifting}

Using limit definition of partial differentiation.

It does work out of the box using MathComp )in that limits exist but does require more lemmas as it is a new library.
Just a lot more detail on this proof when finished.

\section{Other notes}

Complexity increase depending on DL is. While "complexity" is not easy to measure for a DL on paper it is obvious in its impact in formalisation. Associativity of conjunction for Łukawsiewic, Godel and product all have around 10 lines (get exact after cleanup) - but Yager takes 38 (again, correct after cleanup). If I prove STL mention that (if case split in analogous ways to other proofs we get 38 cases)

A lot of difficulty is added due to using nary operations for the sake of generality (bigops, etc.). Will need to have good reasoning for that (primarily htat it's following the very general idea behind LDL).

\bibliographystyle{plainnat}
\bibliography{bibliography}
\end{document}

### 2023-12-20 ITP draft/draft.tex
related work & introduction
some citations missing

@@
-\documentclass[]{article}
+\documentclass[a4paper,10pt]{article}
 \usepackage{amsfonts}
@@
 \usepackage{adjustbox}
+
+\usepackage[numbers]{natbib}
 
@@
 \begin{abstract}
-
 \end{abstract}
@@
 
+Using formal methods, such as using proof assistants or abstract interpretation, in combination with machine learning (ML) is a relatively new approach that has been gaining popularity in recent years \cite{urban2021review,9842406}.
+This can be connected primarily to the rising importance of machine learning verification and an increased awareness of the fragility of machine learning models, both of which value the guarantees that formal methods can provide CITATION.
+Traditionally confined to the realm of software and hardware verification, formal methods are increasingly being repurposed to scrutinize the inherently probabilistic nature of ML models.
+
+The integration of formal methods with machine learning poses many challenges, stemming from the inherent differences between the two domains, from the black-box nature of many ML models, through scalability issues especially in the case of neural networks (NNs) to the lack of well-defined specifications required in formal methods.
+It is the last of those that is of most interest to this work.
+While there has been a move toward increasingly rigorous approaches in specification, both for verification and training, such specification are often mathematically complex.
+
+This complexity is inherent especially to neural networks but is widely known to pose a challenge to formalisation tools.
+The most prominent reasons behind this issue come from the necessity to reason about continuous real space, vectors and vector operations as well as the ability to express mathematical concepts such as partial differentiation.
+This is unavoidable when reasoning about certain methods, as due to common usage of gradient descent algorithm in neural network training CITE.
+
+However we believe that there exist current state of the art formalisation tools which are expressive enough to reason about mathematical concepts needed for expressing properties and methods for machine learning.
+
+Mathematical Components (MathComp)~\cite{mathcomp} in Coq CITE stands as a testament to the expressive power of formal proof systems in addressing advanced mathematical concepts.
+At its core, MathComp leverages the Calculus of Inductive Constructions (CIC), the underlying type theory of Coq, to encode mathematical abstractions in a way that allows bridging the gap between mathematics and the formalization of machine learning concepts.
+
+One of such concepts is Logic of Differentiable Logics (LDL), a rigorously defined meta-language used for property based training of neural networks. ADD A MORE IN DEPTH EXPLANATION OF LDL ITSELF
+
+The goal of this paper is two-fold.
+Firstly, we would like to offer it as demonstration that current state of the art already allows for application of formal approaches to well-defined problems in machine learning and is expressive enough to handle the needed underlying mathematical theory.
+To demonstrate this we formalise the meta-language LDL, a method used to train neural networks, together with its properties and proofs, with emphasis on properties desirable in machine learning.
+
+Furthermore we believe that using formal methods approaches can be beneficial not only in verification, but also in helping to deepen the understanding of machine learning as well as enforcing a more formal approach to more mathematical areas of machine learning.
+For future programming languages created for AI verification  proofs may, in the future, become compulsory, enforcing an additional level of rigour~\cite{FoMLAS2023:Vehicle_Tutorial_Neural_Network,seshia2022toward}.
+
+\section{Related work}
+
+To the best of our knowledge there is very little work that is closely related to our approach and involves applying formal methods to the training approaches itself instead of an already trained machine learning model or its properties.
+
+The primary area where formal methods and machine learning combine is verification.
+This field has been rapidly gaining traction in the last few years and there are many different approaches present in neural network verification\cite{urban2021review,9842406}.
+One of them is abstract interpretation, used by many state of the art verifiers CITE PARTICULAR VERIFIER .
+These methods are partial formal verification techniques and are sound but not complete~\cite{easterbrook1998formal}.
+There are however also approaches that are both sound and complete based on different techniques including use of SMTs (Satisfiability Modulo Theory) \cite{katz2019marabou} solvers or MILP (Mixed Integer Linear Programming).
+There has also been similar work done for other machine learning approaches than neural networks such as for support vector machines (SVMs)~\cite{ranzato2019robustness} or Decision Tree Ensembles~\cite{einziger2019verifying} though we do not focus on such models in our work.
+
+However majority of approaches involving formal methods and machine learning  apply to already trained machine learning models, and not approaches such as constraint-based training itself and thus are not very closely related to this work.
+This area of research is much less explored as of yet, primarily due the difficulty of expressing the mathematical theory underlying neural networks.
+There have however been attempts at it such as \cite{chevallier2022constrained} who formalise an embedding of linear temporal logic and an associated evaluation function in Isabelle to prove its soundness - and furthermore are able to produce OCaml versions of the aforementioned functions that can be integrated with Python and PyTorch eliminating many risks associated with ad-hoc implementations.
+While another related example are formalisation of other types of algorithms~\cite{daukantas2021trimming} or formalisation of neural networks using Haskell~\cite{xie2023haskell}.
+They, while not directly related to the subject matter, utilise similar approaches and also have to contend with similar issues related to mathematical complexity.
+
+POSSIBLY AN ADDITIONAL PARAGRAPH ON NON-ML MATH FORMALISATION THAT IS TANGENTIALLY SIMILIAR.
+
+\section{Background - notes}
+Notes on what probably needs to make its way into background (non-exhaustive)
 \begin{itemize}
-	\item formal methods in combination with machine learning as approach
-	\item normally for verification
-	\item use it to deepen understanding/inform ML research itself
-\end{itemize}
-
-\section{Background}
-\begin{itemize}
-	\item A good explanation of DL and similiar methods in machine learning. Including syntax, semantics in tables.
+	\item A good explanation of DL and similar methods in machine learning. Including syntax, semantics in tables. Also explanation of loss functions and their role in training.
 	\item Possibly properties here, not later.
-	\item Other work on similiar/related topics (pointing out most formalise finalised trained models, not methods of training itself so can't find a good comparison)
+	\item Other work on similar/related topics (pointing out most formalise finalised trained models, not methods of training itself so can't find a good comparison)
 \end{itemize}
@@
 
- Formalization of an embedding of linear temporal logic and an associated evaluation function to prove its soundness in Isabelle and producing OCaml versions of the aforementioned functions that can be integrated with Python and PyTorch \cite{chevallier2022constrained}.
-Formalisation of other types of algorithms~\cite{daukantas2021trimming} or formalisation of neural networks using Haskell~\cite{xie2023haskell}
+\section{Temporary Name: Formalisation - notes}
 
- Reviews of works in formal methods in machine learning (not only neural networks) \cite{urban2021review,9842406}. These are however mostly verification work.
-
- Generally most work is done on trained models - not on the method itself - with very few exceptions.
-Need to find more of the exceptions, put in related work with caveat that they are not always very directly related to this.
-
-\section{Temporary Name: Formalisation}
+General notes to self on formalisation.
 

### 2023-12-21 ITP draft/draft.tex
minor polishing

@@
 
-POSSIBLY AN ADDITIONAL PARAGRAPH ON NON-ML MATH FORMALISATION THAT IS TANGENTIALLY SIMILIAR.
+POSSIBLY AN ADDITIONAL PARAGRAPH ON NON-ML MATH FORMALISATION THAT IS TANGENTIALLY SIMILAR.
 
@@
 
-Different functions, lemmas, and things that peraphs stand out.
+Different functions, lemmas, and things that perhaps stand out.
 
@@
 	\item a word on translation of indices and vectors using tuples
-	\item mention Real library from Mathcomp to handle reals
+	\item mention Real library from MathComp to handle reals
 	\item $translation$ - a semantics for all fuzzy DLs.

### 2024-01-03 ITP draft/draft_lipics.tex
reformatting to lipics

--- 2023-12-21 ITP draft/draft.tex
@@
-\documentclass[a4paper,10pt]{article}
-\usepackage{amsfonts}
-\usepackage{amsthm}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{adjustbox}
+\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
 
-\usepackage[numbers]{natbib}
+\bibliographystyle{plainurl}% the mandatory bibstyle
 
-\newcommand{\tempty}[1]{\llbracket #1 \rrbracket}
-\newcommand{\conjM}{\bigwedge\nolimits_M}
-\newcommand{\Real}{\mathbb{R}}
+\title{Placeholder title: Formal methods in constraint-based learning} %TODO Please add
 
-%opening
-\title{Placeholder title: Formal methods in constraint-based learning}
-\author{}
+%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
+
+\author{Jane {Open Access}}{Dummy University Computing Laboratory, [optional: Address], Country \and My second affiliation, Country \and \url{http://www.myhomepage.edu} }{johnqpublic@dummyuni.org}{https://orcid.org/0000-0002-1825-0097}{(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
+Please provide at least the name of the affiliation and the country.
+The full address is optional.
+Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.
+
+%\author{Joan R.
+Public\footnote{Optional footnote, e.g.
+to mark corresponding author}}{Department of Informatics, Dummy College, [optional: Address], Country}{joanrpublic@dummycollege.org}{[orcid]}{[funding]}
+
+\authorrunning{running author} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'
+
+\Copyright{copyright} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+
+\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm
+
+\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords
+
+\category{} %optional, e.g. invited paper
+
+\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
+%\relatedversiondetails[linktext={opt.
+text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g.
+Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional
+
+%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
+%\supplementdetails[linktext={opt.
+text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g.
+Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional
+
+%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors.
+Please enter author specific funding statements as fifth argument of the \author macro.
+
+%\acknowledgements{I want to thank \dots}%optional
+
+%\nolinenumbers %uncomment to disable line numbering
+
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\EventEditors{John Q. Open and Joan R. Access}
+\EventNoEds{2}
+\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
+\EventShortTitle{CVIT 2016}
+\EventAcronym{CVIT}
+\EventYear{2016}
+\EventDate{December 24--27, 2016}
+\EventLocation{Little Whinging, United Kingdom}
+\EventLogo{}
+\SeriesVolume{42}
+\ArticleNo{23}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%TODO mandatory: add short abstract of the document
 \begin{abstract}
+Abstract.
 \end{abstract}
@@
 \section{Background - notes}
+
 Notes on what probably needs to make its way into background (non-exhaustive)
@@
 	\item Possibly properties here, not later.
-	\item Other work on similar/related topics (pointing out most formalise finalised trained models, not methods of training itself so can't find a good comparison)
 \end{itemize}
-
-\textbf{Some of other work (more or less related depending on the specific case:}
 
@@
 
-\bibliographystyle{plainnat}
+%%
+%% Bibliography
+%%
+
+%% Please use bibtex,
+
 \bibliography{bibliography}
+
+\appendix
+
 \end{document}

### 2024-01-03 ITP draft/lipics/lipics-v2021-sample-article.tex
reformatting to lipics

\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Placeholder title: Formal methods in constraint-based learning} %TODO Please add

%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line

\author{Jane {Open Access}}{Dummy University Computing Laboratory, [optional: Address], Country \and My second affiliation, Country \and \url{http://www.myhomepage.edu} }{johnqpublic@dummyuni.org}{https://orcid.org/0000-0002-1825-0097}{(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Joan R. Public\footnote{Optional footnote, e.g. to mark corresponding author}}{Department of Informatics, Dummy College, [optional: Address], Country}{joanrpublic@dummycollege.org}{[orcid]}{[funding]}

\authorrunning{J. Open Access and J.\,R. Public} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Jane Open Access and Joan R. Public} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords

\category{} %optional, e.g. invited paper

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
%\supplementdetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g. Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

\acknowledgements{I want to thank \dots}%optional

%\nolinenumbers %uncomment to disable line numbering

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Praesent convallis orci arcu, eu mollis dolor. Aliquam eleifend suscipit lacinia. Maecenas quam mi, porta ut lacinia sed, convallis ac dui. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse potenti. 
\end{abstract}

\section{Typesetting instructions -- Summary}
\label{sec:typesetting-summary}

LIPIcs is a series of open access high-quality conference proceedings across all fields in informatics established in cooperation with Schloss Dagstuhl. 
In order to do justice to the high scientific quality of the conferences that publish their proceedings in the LIPIcs series, which is ensured by the thorough review process of the respective events, we believe that LIPIcs proceedings must have an attractive and consistent layout matching the standard of the series.
Moreover, the quality of the metadata, the typesetting and the layout must also meet the requirements of other external parties such as indexing service, DOI registry, funding agencies, among others. The guidelines contained in this document serve as the baseline for the authors, editors, and the publisher to create documents that meet as many different requirements as possible.

Please comply with the following instructions when preparing your article for a LIPIcs proceedings volume. 
\paragraph*{Minimum requirements}

\begin{itemize}
\item Use pdflatex and an up-to-date \LaTeX{} system.
\item Use further \LaTeX{} packages and custom made macros carefully and only if required.
\item Use the provided sectioning macros: \verb+\section+, \verb+\subsection+, \verb+\subsubsection+, \linebreak \verb+\paragraph+, \verb+\paragraph*+, and \verb+\subparagraph*+.
\item Provide suitable graphics of at least 300dpi (preferably in PDF format).
\item Use BibTeX and keep the standard style (\verb+plainurl+) for the bibliography.
\item Please try to keep the warnings log as small as possible. Avoid overfull \verb+\hboxes+ and any kind of warnings/errors with the referenced BibTeX entries.
\item Use a spellchecker to correct typos.
\end{itemize}

\paragraph*{Mandatory metadata macros}
Please set the values of the metadata macros carefully since the information parsed from these macros will be passed to publication servers, catalogues and search engines.
Avoid placing macros inside the metadata macros. The following metadata macros/environments are mandatory:
\begin{itemize}
\item \verb+\title+ and, in case of long titles, \verb+\titlerunning+.
\item \verb+\author+, one for each author, even if two or more authors have the same affiliation.
\item \verb+\authorrunning+ and \verb+\Copyright+ (concatenated author names)\\
The \verb+\author+ macros and the \verb+\Copyright+ macro should contain full author names (especially with regard to the first name), while \verb+\authorrunning+ should contain abbreviated first names.
\item \verb+\ccsdesc+ (ACM classification, see \url{https://www.acm.org/publications/class-2012}).
\item \verb+\keywords+ (a comma-separated list of keywords).
\item \verb+\relatedversion+ (if there is a related version, typically the ``full version''); please make sure to provide a persistent URL, e.\,g., at arXiv.
\item \verb+\begin{abstract}...\end{abstract}+ .
\end{itemize}

\paragraph*{Please do not \ldots} %Do not override the \texttt{\seriesstyle}-defaults}
Generally speaking, please do not override the \texttt{lipics-v2021}-style defaults. To be more specific, a short checklist also used by Dagstuhl Publishing during the final typesetting is given below.
In case of \textbf{non-compliance} with these rules Dagstuhl Publishing will remove the corresponding parts of \LaTeX{} code and \textbf{replace it with the \texttt{lipics-v2021} defaults}. In serious cases, we may reject the LaTeX-source and expect the corresponding author to revise the relevant parts.
\begin{itemize}
\item Do not use a different main font. (For example, the \texttt{times} package is forbidden.)
\item Do not alter the spacing of the \texttt{lipics-v2021.cls} style file.
\item Do not use \verb+enumitem+ and \verb+paralist+. (The \texttt{enumerate} package is preloaded, so you can use
 \verb+\begin{enumerate}[(a)]+ or the like.)
\item Do not use ``self-made'' sectioning commands (e.\,g., \verb+\noindent{\bf My+ \verb+Paragraph}+).
\item Do not hide large text blocks using comments or \verb+\iffalse+ $\ldots$ \verb+\fi+ constructions. 
\item Do not use conditional structures to include/exclude content. Instead, please provide only the content that should be published -- in one file -- and nothing else.
\item Do not wrap figures and tables with text. In particular, the package \texttt{wrapfig} is not supported.
\item Do not change the bibliography style. In particular, do not use author-year citations. (The
\texttt{natbib} package is not supported.)
\end{itemize}

\enlargethispage{\baselineskip}

This is only a summary containing the most relevant details. Please read the complete document ``LIPIcs: Instructions for Authors and the \texttt{lipics-v2021} Class'' for all details and don't hesitate to contact Dagstuhl Publishing (\url{mailto:publishing@dagstuhl.de}) in case of questions or comments:
\href{http://drops.dagstuhl.de/styles/lipics-v2021/lipics-v2021-authors/lipics-v2021-authors-guidelines.pdf}{\texttt{http://drops.dagstuhl.de/styles/lipics-v2021/\newline lipics-v2021-authors/lipics-v2021-authors-guidelines.pdf}}

\section{Lorem ipsum dolor sit amet}

Lorem ipsum dolor sit amet, consectetur adipiscing elit \cite{DBLP:journals/cacm/Knuth74}. Praesent convallis orci arcu, eu mollis dolor. Aliquam eleifend suscipit lacinia. Maecenas quam mi, porta ut lacinia sed, convallis ac dui. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse potenti. Donec eget odio et magna ullamcorper vehicula ut vitae libero. Maecenas lectus nulla, auctor nec varius ac, ultricies et turpis. Pellentesque id ante erat. In hac habitasse platea dictumst. Curabitur a scelerisque odio. Pellentesque elit risus, posuere quis elementum at, pellentesque ut diam. Quisque aliquam libero id mi imperdiet quis convallis turpis eleifend.

\begin{lemma}[Lorem ipsum]
\label{lemma:lorem}
Vestibulum sodales dolor et dui cursus iaculis. Nullam ullamcorper purus vel turpis lobortis eu tempus lorem semper. Proin facilisis gravida rutrum. Etiam sed sollicitudin lorem. Proin pellentesque risus at elit hendrerit pharetra. Integer at turpis varius libero rhoncus fermentum vitae vitae metus.
\end{lemma}

\begin{proof}
Cras purus lorem, pulvinar et fermentum sagittis, suscipit quis magna.

\proofsubparagraph*{Just some paragraph within the proof.}
Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
\begin{claim}
content...
\end{claim}
\begin{claimproof}
content...
    \begin{enumerate}
        \item abc abc abc \claimqedhere{}
    \end{enumerate}
\end{claimproof}

\end{proof}

\begin{corollary}[Curabitur pulvinar, \cite{DBLP:books/mk/GrayR93}]
\label{lemma:curabitur}
Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
\end{corollary}

\begin{proposition}\label{prop1}
This is a proposition
\end{proposition}

\autoref{prop1} and \cref{prop1} \ldots

\subsection{Curabitur dictum felis id sapien}

Curabitur dictum \cref{lemma:curabitur} felis id sapien \autoref{lemma:curabitur} mollis ut venenatis tortor feugiat. Curabitur sed velit diam. Integer aliquam, nunc ac egestas lacinia, nibh est vehicula nibh, ac auctor velit tellus non arcu. Vestibulum lacinia ipsum vitae nisi ultrices eget gravida turpis laoreet. Duis rutrum dapibus ornare. Nulla vehicula vulputate iaculis. Proin a consequat neque. Donec ut rutrum urna. Morbi scelerisque turpis sed elit sagittis eu scelerisque quam condimentum. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nec faucibus leo. Cras ut nisl odio, non tincidunt lorem. Integer purus ligula, venenatis et convallis lacinia, scelerisque at erat. Fusce risus libero, convallis at fermentum in, dignissim sed sem. Ut dapibus orci vitae nisl viverra nec adipiscing tortor condimentum \cite{DBLP:journals/cacm/Dijkstra68a}. Donec non suscipit lorem. Nam sit amet enim vitae nisl accumsan pretium.

\begin{lstlisting}[caption={Useless code.},label=list:8-6,captionpos=t,float,abovecaptionskip=-\medskipamount]
for i:=maxint to 0 do 
begin 
    j:=square(root(i));
end;
\end{lstlisting}

\subsection{Proin ac fermentum augue}

Proin ac fermentum augue. Nullam bibendum enim sollicitudin tellus egestas lacinia euismod orci mollis. Nulla facilisi. Vivamus volutpat venenatis sapien, vitae feugiat arcu fringilla ac. Mauris sapien tortor, sagittis eget auctor at, vulputate pharetra magna. Sed congue, dui nec vulputate convallis, sem nunc adipiscing dui, vel venenatis mauris sem in dui. Praesent a pretium quam. Mauris non mauris sit amet eros rutrum aliquam id ut sapien. Nulla aliquet fringilla sagittis. Pellentesque eu metus posuere nunc tincidunt dignissim in tempor dolor. Nulla cursus aliquet enim. Cras sapien risus, accumsan eu cursus ut, commodo vel velit. Praesent aliquet consectetur ligula, vitae iaculis ligula interdum vel. Integer faucibus faucibus felis.

\begin{itemize}
\item Ut vitae diam augue. 
\item Integer lacus ante, pellentesque sed sollicitudin et, pulvinar adipiscing sem. 
\item Maecenas facilisis, leo quis tincidunt egestas, magna ipsum condimentum orci, vitae facilisis nibh turpis et elit. 
\end{itemize}

\begin{remark}
content...
\end{remark}

\section{Pellentesque quis tortor}

Nec urna malesuada sollicitudin. Nulla facilisi. Vivamus aliquam tempus ligula eget ornare. Praesent eget magna ut turpis mattis cursus. Aliquam vel condimentum orci. Nunc congue, libero in gravida convallis \cite{DBLP:conf/focs/HopcroftPV75}, orci nibh sodales quam, id egestas felis mi nec nisi. Suspendisse tincidunt, est ac vestibulum posuere, justo odio bibendum urna, rutrum bibendum dolor sem nec tellus.

\begin{lemma} [Quisque blandit tempus nunc]
Sed interdum nisl pretium non. Mauris sodales consequat risus vel consectetur. Aliquam erat volutpat. Nunc sed sapien ligula. Proin faucibus sapien luctus nisl feugiat convallis faucibus elit cursus. Nunc vestibulum nunc ac massa pretium pharetra. Nulla facilisis turpis id augue venenatis blandit. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
\end{lemma}

Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.

\section{Morbi eros magna}

Morbi eros magna, vestibulum non posuere non, porta eu quam. Maecenas vitae orci risus, eget imperdiet mauris. Donec massa mauris, pellentesque vel lobortis eu, molestie ac turpis. Sed condimentum convallis dolor, a dignissim est ultrices eu. Donec consectetur volutpat eros, et ornare dui ultricies id. Vivamus eu augue eget dolor euismod ultrices et sit amet nisi. Vivamus malesuada leo ac leo ullamcorper tempor. Donec justo mi, tempor vitae aliquet non, faucibus eu lacus. Donec dictum gravida neque, non porta turpis imperdiet eget. Curabitur quis euismod ligula.

%%
%% Bibliography
%%

%% Please use bibtex,

\bibliography{lipics-v2021-sample-article}

\appendix

\section{Styles of lists, enumerations, and descriptions}\label{sec:itemStyles}

List of different predefined enumeration styles:

\begin{itemize}
\item \verb|\begin{itemize}...\end{itemize}|
\item \dots
\item \dots
%\item \dots
\end{itemize}

\begin{enumerate}
\item \verb|\begin{enumerate}...\end{enumerate}|
\item \dots
\item \dots
%\item \dots
\end{enumerate}

\begin{alphaenumerate}
\item \verb|\begin{alphaenumerate}...\end{alphaenumerate}|
\item \dots
\item \dots
%\item \dots
\end{alphaenumerate}

\begin{romanenumerate}
\item \verb|\begin{romanenumerate}...\end{romanenumerate}|
\item \dots
\item \dots
%\item \dots
\end{romanenumerate}

\begin{bracketenumerate}
\item \verb|\begin{bracketenumerate}...\end{bracketenumerate}|
\item \dots
\item \dots
%\item \dots
\end{bracketenumerate}

\begin{description}
\item[Description 1] \verb|\begin{description} \item[Description 1]  ...\end{description}|
\item[Description 2] Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\item[Description 3]  \dots
%\item \dots
\end{description}

\cref{testenv-proposition} and \autoref{testenv-proposition} ...

\section{Theorem-like environments}\label{sec:theorem-environments}

List of different predefined enumeration styles:

\begin{theorem}\label{testenv-theorem}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{theorem}

\begin{lemma}\label{testenv-lemma}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{lemma}

\begin{corollary}\label{testenv-corollary}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{corollary}

\begin{proposition}\label{testenv-proposition}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{proposition}

\begin{conjecture}\label{testenv-conjecture}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{conjecture}

\begin{observation}\label{testenv-observation}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{observation}

\begin{exercise}\label{testenv-exercise}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{exercise}

\begin{definition}\label{testenv-definition}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{definition}

\begin{example}\label{testenv-example}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{example}

\begin{note}\label{testenv-note}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{note}

\begin{note*}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{note*}

\begin{remark}\label{testenv-remark}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{remark}

\begin{remark*}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{remark*}

\begin{claim}\label{testenv-claim}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{claim}

\begin{claim*}\label{testenv-claim2}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{claim*}

\begin{proof}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{proof}

\begin{claimproof}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{claimproof}

\end{document}

### 2024-01-04 ITP draft/draft_lipics.tex
minor comments and rm auto files

@@
 
+\def\rei#1{{\color{red}[NB(rei):#1]}}
+
 \begin{document}
@@
 Furthermore we believe that using formal methods approaches can be beneficial not only in verification, but also in helping to deepen the understanding of machine learning as well as enforcing a more formal approach to more mathematical areas of machine learning.
-For future programming languages created for AI verification  proofs may, in the future, become compulsory, enforcing an additional level of rigour~\cite{FoMLAS2023:Vehicle_Tutorial_Neural_Network,seshia2022toward}.
+For future programming languages created for AI verification  proofs may, in the future\rei{``future'' appears twice}, become compulsory, enforcing an additional level of rigour~\cite{FoMLAS2023:Vehicle_Tutorial_Neural_Network,seshia2022toward}.
+
+\rei{I suspect that many reviewers look for a ``our contributions'' paragraph}
 
@@
 This area of research is much less explored as of yet, primarily due the difficulty of expressing the mathematical theory underlying neural networks.
-There have however been attempts at it such as \cite{chevallier2022constrained} who formalise an embedding of linear temporal logic and an associated evaluation function in Isabelle to prove its soundness - and furthermore are able to produce OCaml versions of the aforementioned functions that can be integrated with Python and PyTorch eliminating many risks associated with ad-hoc implementations.
-While another related example are formalisation of other types of algorithms~\cite{daukantas2021trimming} or formalisation of neural networks using Haskell~\cite{xie2023haskell}.
+There have however been attempts at it such as \cite{chevallier2022constrained} who formalise an embedding of linear temporal logic and an associated evaluation function in Isabelle to prove its soundness---and furthermore are able to produce OCaml versions of the aforementioned functions that can be integrated with Python and PyTorch eliminating many risks associated with ad-hoc implementations.
+While another related example are\rei{singular?} formalisation of other types of algorithms~\cite{daukantas2021trimming} or formalisation of neural networks using Haskell~\cite{xie2023haskell}.
 They, while not directly related to the subject matter, utilise similar approaches and also have to contend with similar issues related to mathematical complexity.
@@
 On associativity and commutativity for the DLs that have them (and that associativity is much more tricky then commutativity to prove).
-Not entirely sure if antyhing more interesting is here specifically so likely not its own subsection.
+Not entirely sure if anything more interesting is here specifically so likely not its own subsection.
 
@@
 
-It does work out of the box using MathComp )in that limits exist but does require more lemmas as it is a new library.
+It does work out of the box using MathComp in that limits exist but does require more lemmas as it is a new library.
 Just a lot more detail on this proof when finished.
@@
 Complexity increase depending on DL is.
-While "complexity" is not easy to measure for a DL on paper it is obvious in its impact in formalisation.
+While ``complexity'' is not easy to measure for a DL on paper it is obvious in its impact in formalisation.
 Associativity of conjunction for Łukawsiewic, Godel and product all have around 10 lines (get exact after cleanup) - but Yager takes 38 (again, correct after cleanup).

### 2024-01-09 ITP draft/draft_lipics.tex
assoc + idemp + shadow lift
redoing intro/related work

--- 2024-01-03 ITP draft/draft_lipics.tex
@@
 
+\usepackage{tcolorbox}
+\usepackage{local-macros}
+\usepackage{syntax}
 \bibliographystyle{plainurl}% the mandatory bibstyle
@@
 
+%COMMENT MACROS
+\newtcolorbox{authorComment}[1]{colback=#1}
+\newcommand{\natalia}[1]{\begin{authorComment}{red!15}Natalia: #1\end{authorComment}}
+
 \begin{document}
@@
 
-Using formal methods, such as using proof assistants or abstract interpretation, in combination with machine learning (ML) is a relatively new approach that has been gaining popularity in recent years \cite{urban2021review,9842406}.
-This can be connected primarily to the rising importance of machine learning verification and an increased awareness of the fragility of machine learning models, both of which value the guarantees that formal methods can provide CITATION.
-Traditionally confined to the realm of software and hardware verification, formal methods are increasingly being repurposed to scrutinize the inherently probabilistic nature of ML models.
+Using Interactive Theorem Provers (ITPs) in combination with machine learning (ML) is a very new approach dating only a few years back CITE. 
+The integration of formal methods with machine learning poses many challenges, stemming from the inherent differences between the two domains, from the black-box nature of many ML models, through scalability issues especially in the case of neural networks (NNs) to the lack of well-defined specifications required in formal methods.
 
-The integration of formal methods with machine learning poses many challenges, stemming from the inherent differences between the two domains, from the black-box nature of many ML models, through scalability issues especially in the case of neural networks (NNs) to the lack of well-defined specifications required in formal methods.
-It is the last of those that is of most interest to this work.
-While there has been a move toward increasingly rigorous approaches in specification, both for verification and training, such specification are often mathematically complex.
+However in the recent years there has been a growing interest in the intersection of these two fields.
+One of the first attempts to use Coq in conjunction with machine learning is MLCert~\cite{bagnall2019certifying}, a system for mechanized proofs of the generalisation bounds of an ML model in Coq, providing bounds on expected error.
+MLCert is compatible with deep learning frameworks such as TensorFlow but the classifier does need to be encoded by hand.
+Coq can be easier used for reasoning about well defined smaller machine learning algorithms, for example for proving convergence of a one-layer perceptron in specific conditions~\cite{murphy2017verified}.
 
-This complexity is inherent especially to neural networks but is widely known to pose a challenge to formalisation tools.
-The most prominent reasons behind this issue come from the necessity to reason about continuous real space, vectors and vector operations as well as the ability to express mathematical concepts such as partial differentiation.
-This is unavoidable when reasoning about certain methods, as due to common usage of gradient descent algorithm in neural network training CITE.
+The existing work is not limited to Coq.
+It has been demonstrated that a neural network can be encoded in a theorem prover, in this case Isabelle/HOL, in order to prove its safety \cite{brucker2023verifying}.
+Vehicle, a tool for enforcing specifications on neural networks, is also capable of exporting a verified specification to Agda to enable reasoning about the network when considering safety of a larger system\cite{daggitt2023compiling}.
 
-However we believe that there exist current state of the art formalisation tools which are expressive enough to reason about mathematical concepts needed for expressing properties and methods for machine learning.
+The existing attempts at combining ITPs and machine learning have encountered some of the same problems that originate form the inherent nature of machine learning as a field and are challenging for ITPs.
+Among the most theoretical ones are the necessity to deal with real domain as well as vectors in order to reason about data as well as need for support for linear algebra and certain areas of calculus.
+This mathematical complexity is inherent especially to neural networks but is widely known to pose a challenge to formalisation tools.
+Lastly there is also an issue of scale when formalising large models.
 
-Mathematical Components (MathComp)~\cite{mathcomp} in Coq CITE stands as a testament to the expressive power of formal proof systems in addressing advanced mathematical concepts.
+There are many aspects of neural networks that can be formalised. \natalia{Need to find a nice transition in this paragraph}In particular one we investigate in this work is differentiable logics.
+
+Differentiable logics belong to a method of property-based training of neural networks.
+The idea behind differentiable logic is to use a property that is desirable in a neural network and use it to generate a custom property based loss function.
+A standard \emph{loss function} $\losssymbol: \Real^n \times \Real^m \rightarrow \Real$ given a pair $(\x, \y)$ of input and desired output, a neural network $f$, calculates how much $f(\x)$ differs from the desired output $\y$ \cite{wang2020comprehensive}.
+
+A property based loss will instead aim to penalise the network for deviating from said property.
+Training a neural network with a combination of a property based loss function and a standard data-driven one will result in a network trained to fit both the data and the property.
+
+A DL itself consists of syntax - the logical language used to express properties - and semantics - the "translation" from the syntax to a loss function.
+A general typed syntax and semantics for expressing different popular DLs that we use as a reference point is Logic of Differentiable Logics (LDL) CITE.
+The syntax of LDL is a typed first order logic language with addition of lambdas however in this paper we will consider its propositional fragment which can be seen in Figure \ref{fig:syntax}.
+Importantly the type system includes the existance of vectors, vector indicies and real numbers.
+
+The syntax of LDL is then interpreted into a loss function.
+The semantics are parametric based on the choice of interpretation for comparisons and logical connectives which differ between various differentiable logics.
+
+\natalia{Let me know if I should expand on LDL here or in the background section instead.
+Running example might be good to illustrate it since I think this explanation alone is likely confusing but I don't know what sort of running example would be compelling for ITP audience}
+
+\begin{figure}[t]
+	\small
+	%\footnotesize
+	\centering
+	\begin{subfigure}[t]{0.45\textwidth}
+		\begin{grammar}
+			<expr> $\ni e$ ::=  
+			$\identClass$ | $\netClass$ | \elReal $\: \in \Real$ | \elNat $\: \in \Nat$ | \elBool $\: \in \Bool$
+			\alt \App{\exprClass}{\exprClass} | \Lam{\identClass}{\typeClass}{\exprClass} | \Let{\identClass}{\typeClass}{\exprClass}{\exprClass}
+			\alt \AndSymbol{} | \OrSymbol{} | \NotSymbol{} | \ImpliesSymbol{} | \AddSymbol{} | \NegSymbol{} | \MulSymbol{}
+			\alt \NeqSymbol{} | \LeqSymbol{} | \GeqSymbol{} | \LeSymbol{} | \GeSymbol{} | \EqSymbol{}
+			\alt \Seq{\exprClass}{\exprClass} | \AtSymbol{}
+%			\alt \Forall{\identClass}{\typeClass}{\exprClass} | \Exists{\identClass}{\typeClass}{\exprClass}
+
+		\end{grammar}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.5\textwidth}
+		\begin{grammar}
+			<type> $\ni \tau$  ::= 
+			\FunType{s}{\typeClass} 
+			| s
+
+			<simple type> $\ni s$ ::= 
+			\alt \BoolType
+			\alt \RealType
+			\alt \VecType{n}   |  \FinType{n}  \hspace{3em} for $n \in \Nat$
+		\end{grammar}
+	\end{subfigure}
+	\setlength{\belowcaptionskip}{-15pt} 
+	\caption{Specification Language of LDL: expressions and types. % (e.g. $x \leq y$ is syntactic sugar for $\leq x\ y$).
+	}
+	\label{fig:syntax}
+\end{figure}
+
+%Furthermore we believe that using formal methods approaches can be beneficial not only in verification, but also in helping to deepen the understanding of machine learning as well as enforcing a more formal approach to more mathematical areas of machine learning.
+%For future programming languages created for AI verification  proofs may, in the future, become compulsory, enforcing an additional level of rigour~\cite{FoMLAS2023:Vehicle_Tutorial_Neural_Network,seshia2022toward}.
+
+The problem that we address is that DL formalisation papers are often unclear or wrong in claims about certain properties of DLs they claim CITE MYSELF, STL.
+The approach we use is formalisation of the the meta-language LDL in the Coq proof assistant.
+While as mentioned the mathematical context needed to reason about DLs is challenging for theorem provers, we believe that there exist current state of the art formalisation tools which are expressive enough and we use paper as demonstration that current state of the art already allows for application of formal approaches to well-defined problems in machine learning.
+
+Hence we have chosen to use Mathematical Components (MathComp)~\cite{mathcomp} in Coq CITE which stands as a testament to the expressive power of formal proof systems in addressing advanced mathematical concepts.
 At its core, MathComp leverages the Calculus of Inductive Constructions (CIC), the underlying type theory of Coq, to encode mathematical abstractions in a way that allows bridging the gap between mathematics and the formalization of machine learning concepts.
 
-One of such concepts is Logic of Differentiable Logics (LDL), a rigorously defined meta-language used for property based training of neural networks. ADD A MORE IN DEPTH EXPLANATION OF LDL ITSELF
+Our first contribution is the formalisation of several best known DLs using Coq and MathComp.
+Our formalisation of the LDL framework is designed to be generic and high-level so that it allows for easy extension to other DLs.
+EXPAND
 
-The goal of this paper is two-fold.
-Firstly, we would like to offer it as demonstration that current state of the art already allows for application of formal approaches to well-defined problems in machine learning and is expressive enough to handle the needed underlying mathematical theory.
-To demonstrate this we formalise the meta-language LDL, a method used to train neural networks, together with its properties and proofs, with emphasis on properties desirable in machine learning.
+Secondly, our implementation of the syntax is non-trivial.
+It utilises dependent types interestingly.
+We have discovered some unintended behaviours in Coq implementation of dependent types which caused standard tactics to fail in proofs.
+\natalia{Should I explain the problem with R as context vs variable here, or in some of the later chapters?}
 
-Furthermore we believe that using formal methods approaches can be beneficial not only in verification, but also in helping to deepen the understanding of machine learning as well as enforcing a more formal approach to more mathematical areas of machine learning.
-For future programming languages created for AI verification  proofs may, in the future, become compulsory, enforcing an additional level of rigour~\cite{FoMLAS2023:Vehicle_Tutorial_Neural_Network,seshia2022toward}.
+Another of our contributions is demonstrating the capabilities of MathComp, including latest additions to the library CITE, when applied to machine learning.
+This is the first successful attempt at formalisation of the LDL calculus which includes dealing with concepts known to be problematic in formalisation and yet prevalent in machine learning such as working in the real domain and dealing with vectors.
+
+Lastly we provide formal proofs of theorems that were not rigorously proven in ML literature CITE MYSELF ALSO STL.
+This refers to both theorems that were stated incorrectly, not listing all of the necessary assumptions as well as incorrect claims about certain DLs holding specific desirable mathematical properties.
+We would like this work to inform design choices in DLs and help fix preexisting issues.
+We believe that using formal methods approaches can be beneficial not only in verification, but also in helping to deepen the understanding of machine learning as well as enforcing a more formal approach to more mathematical areas of machine learning.
+\natalia{Should I go into detail here what the actual issue was or should that be left for later?
+Currently plan to discuss the STL paper shadow-lifting/assoc/idempotence example in later chapter as per plan}
 
@@
 
-To the best of our knowledge there is very little work that is closely related to our approach and involves applying formal methods to the training approaches itself instead of an already trained machine learning model or its properties.
+To the best of our knowledge there is very little work that is closely related to our approach and involves applying formal methods and ITPs to the training approaches itself instead of an already trained machine learning model or its properties.
+Furthermore majority of approaches involving formal methods and machine learning  apply to already trained machine learning models, and not approaches such as constraint-based training itself and thus are not very closely related to this work.
+This area of research is much less explored as of yet, primarily due the difficulty of expressing the mathematical theory underlying neural networks.
 
-The primary area where formal methods and machine learning combine is verification.
-This field has been rapidly gaining traction in the last few years and there are many different approaches present in neural network verification\cite{urban2021review,9842406}.
-One of them is abstract interpretation, used by many state of the art verifiers CITE PARTICULAR VERIFIER .
-These methods are partial formal verification techniques and are sound but not complete~\cite{easterbrook1998formal}.
-There are however also approaches that are both sound and complete based on different techniques including use of SMTs (Satisfiability Modulo Theory) \cite{katz2019marabou} solvers or MILP (Mixed Integer Linear Programming).
-There has also been similar work done for other machine learning approaches than neural networks such as for support vector machines (SVMs)~\cite{ranzato2019robustness} or Decision Tree Ensembles~\cite{einziger2019verifying} though we do not focus on such models in our work.
+Relevant work dealing with already trained models does involve encoding trained networks in ITPs, such as Isabelle/HOL \cite{brucker2023verifying}.
+This Isabelle prototype supports importing feedforward neural networks from TensorFlow, eliminating the need to embed each network by hand and it is the first formalisation of neural networks in an interactive theorem prover.
+Similiarly another recent work to set up other interactive theorem provers for verification of neural network by formalising piecewise affine activation functions in Coq~\cite{aleksandrov2023formalizing}.
 
-However majority of approaches involving formal methods and machine learning  apply to already trained machine learning models, and not approaches such as constraint-based training itself and thus are not very closely related to this work.
-This area of research is much less explored as of yet, primarily due the difficulty of expressing the mathematical theory underlying neural networks.
-There have however been attempts at it such as \cite{chevallier2022constrained} who formalise an embedding of linear temporal logic and an associated evaluation function in Isabelle to prove its soundness - and furthermore are able to produce OCaml versions of the aforementioned functions that can be integrated with Python and PyTorch eliminating many risks associated with ad-hoc implementations.
-While another related example are formalisation of other types of algorithms~\cite{daukantas2021trimming} or formalisation of neural networks using Haskell~\cite{xie2023haskell}.
+Other works focus on different properties important to the machine learning community such as generalisation, providing formal guarantees of the degree to which the train network will generalise to new data~\cite{bagnall2019certifying}.
+Since scale and lack of precise mathemtaical definitions are some of the main issues for theorem provers, reasoning about properties of certain smaller but well defined machine learning algorithms lend themselves better to formalisation.
+We can take as an example proving convergance of a as single-layer perceptron~\cite{murphy2017verified}.
+
+While they do not directly formalise networks, there has also been some research related to algorithms, instead of trained models such as formalisation of an embedding of linear temporal logic and an associated evaluation function in Isabelle to prove its soundness - and furthermore are able to produce OCaml versions of the aforementioned functions that can be integrated with Python and PyTorch eliminating many risks associated with ad-hoc implementations\cite{chevallier2022constrained}.
+Another related example are formalisation of other types of algorithms~\cite{daukantas2021trimming} or formalisation of neural networks using Haskell~\cite{xie2023haskell}.
 They, while not directly related to the subject matter, utilise similar approaches and also have to contend with similar issues related to mathematical complexity.
 
-POSSIBLY AN ADDITIONAL PARAGRAPH ON NON-ML MATH FORMALISATION THAT IS TANGENTIALLY SIMILAR.
+Lastly, a larger but less directly related area area where more broadly understood formal methods and machine learning combine is verification.
+This field has been rapidly gaining traction in the last few years and there are many different approaches present in neural network verification\cite{urban2021review,9842406,easterbrook1998formal} including abstract interpretation, SMTs (Satisfiability Modulo Theory) \cite{katz2019marabou} solvers or MILP (Mixed Integer Linear Programming).
+There has also been similar work done for other machine learning approaches than neural networks such as for support vector machines (SVMs)~\cite{ranzato2019robustness} or Decision Tree Ensembles~\cite{einziger2019verifying} though we do not focus on such models in our work.
 
@@
 While "complexity" is not easy to measure for a DL on paper it is obvious in its impact in formalisation.
-Associativity of conjunction for Łukawsiewic, Godel and product all have around 10 lines (get exact after cleanup) - but Yager takes 38 (again, correct after cleanup).
+Associativity of conjunction for Łukawsiewicz, Godel and product all have around 10 lines (get exact after cleanup) - but Yager takes 38 (again, correct after cleanup).
 If I prove STL mention that (if case split in analogous ways to other proofs we get 38 cases)
@@
 Will need to have good reasoning for that (primarily htat it's following the very general idea behind LDL).
+
+\section{Placeholder title: stl example}
 

### 2024-01-09 ITP draft/draft_lipics.tex

@@
 \newcommand{\natalia}[1]{\begin{authorComment}{red!15}Natalia: #1\end{authorComment}}
+
+\def\rei#1{{\color{red}[NB(rei):#1]}}
 
@@
 Complexity increase depending on DL is.
-While "complexity" is not easy to measure for a DL on paper it is obvious in its impact in formalisation.
-Associativity of conjunction for Łukawsiewicz, Godel and product all have around 10 lines (get exact after cleanup) - but Yager takes 38 (again, correct after cleanup).
+While ``complexity'' is not easy to measure for a DL on paper it is obvious in its impact in formalisation.
+Associativity of conjunction for Łukawsiewic, Godel and product all have around 10 lines (get exact after cleanup) - but Yager takes 38 (again, correct after cleanup).
 If I prove STL mention that (if case split in analogous ways to other proofs we get 38 cases)
@@
 
-\section{Placeholder title: stl example}
+\section{Shadow lifting, idempotence and associativity}
+
+The claim that shadow lifting and associativity cannot hold together is inaccurate, we need idempotence too.
+
+To make this formal, let us assume a generic semantic interpretation of and (f), with the following properties.
+
+[[ p /\ q ]] = f [[ p ]]  [[ q ]] (it's structural)
+
+I: [[p /\ p]] = [[p]] (has idempotence)
+
+A: [[p /\ (q /\ r)]] = [[(p /\ q) /\ r]] (has associativity)
+
+(has shadow lifting)
+SL1: d [[p /\ q]] / d p > 0
+SL2: d [[p /\ q]] / d q > 0
+
+[[ 1 /\ (1 /\ 1.1) ]] =? [[ 1 /\ (1.1 /\ 1.1) ]]
+
+[[ 1 /\ (1 /\ 1.1) ]] =A [[ (1 /\ 1) /\ 1.1 ]] =I [[ 1 /\ 1.1 ]] =I [[ 1 /\ (1.1 /\ 1.1) ]]
+
+[[ 1 /\ 1.1 ]] <SL1 [[ 1.1 /\ 1.1 ]]
+->
+[[ 1 /\ (1 /\ 1.1) ]] <SL2 [[ 1 /\ (1.1 /\ 1.1) ]]
 

### 2024-01-09 ITP draft/draft_lipics.tex
gitignore for latex + comment

@@
 
-A DL itself consists of syntax - the logical language used to express properties - and semantics - the "translation" from the syntax to a loss function.
+A DL itself consists of syntax - the logical language used to express properties - and semantics - the ``translation'' from the syntax to a loss function.
 A general typed syntax and semantics for expressing different popular DLs that we use as a reference point is Logic of Differentiable Logics (LDL) CITE.
@@
 At its core, MathComp leverages the Calculus of Inductive Constructions (CIC), the underlying type theory of Coq, to encode mathematical abstractions in a way that allows bridging the gap between mathematics and the formalization of machine learning concepts.
+
+\natalia{I'm not sure if I should be using paragraphs or subsections or what in this style? I'd normally use a paragraph for contributions but they seemed to be numbered by default}
 

### 2024-01-10 ITP draft/draft_lipics.tex
sime comments on the Intro

@@
 
+\usepackage[textwidth=2.5cm,textsize=footnotesize,colorinlistoftodos]{todonotes}
+\newcommand{\knote}[2][]{\todo[inline,color=blue!10,#1]{KATYA: #2}}
+
 \begin{document}
@@
 
-Using Interactive Theorem Provers (ITPs) in combination with machine learning (ML) is a very new approach dating only a few years back CITE. 
-The integration of formal methods with machine learning poses many challenges, stemming from the inherent differences between the two domains, from the black-box nature of many ML models, through scalability issues especially in the case of neural networks (NNs) to the lack of well-defined specifications required in formal methods.
+Using Interactive Theorem Provers (ITPs) in combination with machine learning (ML)
+\knote{``in combination" is too broad; as there were works using Ml to speed up proofs. You only want to discuss using ITP in formalisation of ML/probabilistic reasoning}
+ is a very new approach dating only a few years back CITE.
+ \knote{``new" is to weak, new does not mean useful. So, probably the first sentence should start with motivating why it is a good idea to formalise these things in ITP} 
+The integration of formal methods with machine learning poses many challenges, 
+\knote{This changes focus from ITP to formal methods -- do you really need this?}
+stemming from the inherent differences between the two domains, from the black-box nature of many ML models, through scalability issues especially in the case of neural networks (NNs) to the lack of well-defined specifications required in formal methods.
+
+\knote{I don't know...
+There are choices how to start, but generally, we start with an idea that is well accepted by the community we address to a problem that we assume the audience will accept as interesting.}
+\knote{For this reason, I would go as follows}
+\knote{Formalisation of mathematics has been one of major research topics within the Interactive Theorem Proving (ITP) communities, from early successes in formaisation of the Kepler conjecture~\cite{}, to Four-colour theorem~\cite{}, Odd-Order Theorem~\cite{} and more recently, to consolidated efforts to create comprehensive mathematical libraries in Coq (Mathematical Components~\cite{}) and Lean (lookup names ~\cite{}).
+As statistical machine learning rests on differentiation, probability theory and linear algebra, it was only a matter of time to ask whether there is a strong case for, or specific interest in, formalising machine learning in ITP.
+The first attempt at ITP formalisation of neural networks was made by~\cite{bagnall2019certifying}, shortly followed by analysis
+of different perceptron forms in \cite{de2022use}, both in Coq. In both cases the formalisation aimed at
+verifying abstract properties of neural networks: generalisation bounds in ~\cite{bagnall2019certifying} and various forms of convergent behaviour in \cite{de2022use}.  
+More recent Coq formalisations, e.g.
+by \cite{AleksandrovV23} target proofs of theoretical properties of neural networks seen as functions.
+Say something about work of Alessandro?
+someone in Lean?
+someone in Isabelle?}
+\knote{That was idea 1. General introduction}
+\knote{Idea 2: you need to bring here some motivation why someone wants to study/formalise DL...
+I'd try to explain general motivation, without formally defining loss functions...
+(because defining loss functions implies defining neural nets as well, and you loose the point}
 
